---
title: "Dose_Dependence"
output: html_document
---
```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(lubridate)
```

#dose_dependence dataset
```{r}
dose_dependence<-read_csv("doi_10_25338_B8BS75__v20210727/dose_dependence_data.csv")%>%
  clean_names()
```

```{r}
glimpse(dose_dependence)
```
```{r}
summary(dose_dependence)
```

- actually microbe treatment and initial OD is the same but only the initial OD make the treatment into a number
#graph of mircriobe treatment vs. germinated 
-examine how is the proportion germinated change for each time point and microbetreatment
- undiluted appears to have higher germination <=3h
  - after 3h ---> competition?
  - burst?
- 1:20 have higher germination >3h
```{r}
dose_dependence%>%
  mutate(timepoint=factor(timepoint,
         levels=c("15m","45m","90m","3h","4h","6h","12h","24h")))%>%
  ggplot(aes(x=microbetreatment,
             y=proportion_germinated,
             fill=microbetreatment))+
  geom_boxplot()+
  facet_wrap(~timepoint, nrow=2)+
  theme_minimal()+
  labs(title="Concentration of A.pollinisSCC477 and Proportion Germinated",
       x="Microbe Treatment",
       y="Mean Proportion Germinated")+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```
- undiluted appears to maintained its position in proportion of germinated if using allgerm(The proportion of pollen grains that germinated, including both "number.germ" and "number.tip.burst")
-maybe many pollen quickly being burst
- might indicated higher concentration increase both bursting and germination
```{r}
dose_dependence %>%
  mutate(timepoint=factor(timepoint,
         levels=c("15m","45m","90m","3h","4h","6h","12h","24h")))%>%
  ggplot(aes(x=microbetreatment,
             y=allgerm,
             fill=microbetreatment))+
  geom_boxplot()+
  facet_wrap(~timepoint, nrow=2)+
  labs(title="Concentration of A.pollinisSCC477 and All Germinated",
       x="Microbe Treatment",
       y="Proportion of Germinated")+
  theme_minimal()+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```

#whether treatment accelerates germination
- make time point to same unit
```{r}
dose_dependence <- dose_dependence %>%
  mutate(time_minutes=case_when(
    timepoint=="15m" ~ 15,
           timepoint=="45m" ~ 45,
           timepoint=="90m" ~ 90,
           timepoint=="3h"  ~ 180,
           timepoint=="4h"  ~ 240,
           timepoint=="6h"  ~ 360,
           timepoint=="12h" ~ 720,
           timepoint=="24h" ~ 1440))
```

- except sterile BK all other dilution appears to come to a constant
- maximum germination after a long time?
- Peak germination appears highest for 1:20
- Final/long-term germination appears highest for 1:2
```{r}
dose_dependence%>%
  group_by(microbetreatment, time_minutes)%>%
  summarize(mean_germ=mean(allgerm, na.rm=TRUE)) %>%
  ggplot(aes(x=time_minutes, 
             y=mean_germ,
             color=microbetreatment))+
  geom_point()+
  geom_line()+
  labs(
    title="Relationship Between Concentration of A.pollinisSCC477 and Mean Germinated",
    x="Time point(minute)",
    y="Mean Germinated")+
  theme_minimal()
```

#graph of mircriobe treatment vs. brust 
- longer time --> more burst
- higher concentration ---> more bursting
```{r}
dose_dependence%>%
  mutate(timepoint=factor(timepoint,
         levels=c("15m","45m","90m","3h","4h","6h","12h","24h")))%>%
  ggplot(aes(x=microbetreatment,
             y=allburst,
             fill=microbetreatment))+
  geom_boxplot()+
  facet_wrap(~timepoint, nrow=2)+
  theme_minimal()+
  labs(title="Concentration of A.pollinisSCC477 and Proportion Burst",
       x="Microbe Treatment",
       y="Mean Proportion Germinated")+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```

#graph of initial OD vs. germination
- removed sterile BK because its od is 0
- repetitive with above graph
```{r}
#dose_dependence%>%
#  filter(microbetreatment!="sterile BK") %>%
#  mutate(timepoint=factor(timepoint,
#         levels=c("15m","45m","90m","3h","4h","6h","12h","24h"))) %>%
#  pivot_longer(
#    cols=c(allgerm, allburst),
#    names_to="response",
#    values_to="value")%>%
#  ggplot(aes(x = factor(initial_od), y = value)) +
#  geom_boxplot(fill = "lightblue") +
#  facet_grid(response ~ timepoint) +
#  theme_minimal() +
#  labs(
#    title="Relationship Between Initial OD and Germination/Bursting",
#    x ="Initial OD",
#    y ="Proportion")+
# theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#relationship between brust and germination
-not so useful graph--> don't use
```{r}
#dose_dependence%>%
#  mutate(timepoint=factor(timepoint,
#         levels=c("15m","45m","90m","3h","4h","6h","12h","24h")))%>%
#  ggplot(aes(x=allgerm,
#             y=allburst,
#             color=microbetreatment))+
#  geom_point()+
#  geom_smooth(method=lm, se=F)+
#  facet_wrap(~timepoint, nrow=2)+
#  theme(legend.position="bottom")+
#  labs(
#    title="Relationship between Germination and Bursting",
#    x="Proportion Germinated",
#    y="Proportion Burst",
#    color="Microbe Treatment")
```

