---
title: "Pollen_Microbe"
output: html_document
---
```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(lubridate)
```

#pollen_micriobe dataset
```{r}
pollen_micriobe<-read_csv("doi_10_25338_B8BS75__v20210727/pollen_microbe_data.csv")%>%
  clean_names()
```

```{r}
glimpse(pollen_micriobe)
```
```{r}
summary(pollen_micriobe)
```

#distribution of microbe group
- examine the distribution on each sample
- find does one group having more samples than the other
```{r}
pollen_micriobe %>%
  ggplot(aes(x=microbe_group, fill=microbe_group))+
  geom_bar()+
  theme_minimal()+
  labs(
    title="Distribution of Microbe Groups",
    x=NULL,
    y="Count")+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```

#microbe group and all germination
- compare the germination of all microbe group
- (according to the README allgerm; proportion of pollen grains that germinated, including both "number.germ" and "number.tip.burst )
- proportion germinated is calculated by number.germ/Total.pollen
- include number tip burst give us more accurate overlooks on pollen initiated germination
```{r}
pollen_micriobe%>%
  ggplot(aes(x=microbe_group, 
             y=allgerm, 
             fill=microbe_group))+
  geom_boxplot()+
  facet_wrap(~timepoint)+
  theme_minimal()+
  labs(
    title="All Germination of Different Microbe Group by Timepoints",
    x=NULL,
    y="All Germination")+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```

#microbe group and all germination(distribution)
- examine how germination of each microbe group distributed
- Acinetobacter; more variety in data
```{r}
pollen_micriobe%>%
  ggplot(aes(x=prop_germ))+
  geom_histogram(
    binwidth = 0.05,
    fill="lightblue",
    color="black")+
  theme_minimal()+
  facet_wrap(~microbe_group)+
  labs(
    title="Distribution of Proportion Germinated",
    x="Proportion Germinated",
    y="Count")
```

#microbe group and all  burst
- allburst; The proportion of pollen grains that expelled protoplasm into solution, including both "number.burst" and "number.tip.burst
- which given pollen that burst/pollen released nutrients
- by group the mean burst of Acinetobacter across all different time point is higher than other group
```{r}
pollen_micriobe%>%
  ggplot(aes(x=microbe_group, y=allburst, fill=microbe_group))+
  geom_boxplot()+
  facet_wrap(~timepoint)+
  theme_minimal()+
  labs(
    title="All Burst of Different Microbe Group by Timepoints",
    x=NULL,
    y="Porprotion Burst")+
  theme(legend.position="none",
        axis.text.x=element_text(angle=45, hjust=1))
```

#microbe group and all burst(distribution)
- more burst in Acintebacter
```{r}
pollen_micriobe%>%
  ggplot(aes(x=prop_burst))+
  geom_histogram(
    binwidth=0.05,
    fill="blue",
    color="black",
    alpha=0.6)+
  theme_minimal()+
  facet_wrap(~microbe_group)+
  labs(
    title="Distribution of Proportion Burst",
    x="Proportion Burst",
    y="Count")
```
#count timespot
```{r}
pollen_micriobe%>%
  count(timepoint)
```

#germination and time
- check if longer time did cause larger germination ---> does show
```{r}
pollen_micriobe%>%
  group_by(microbe_group, timepoint)%>%
  summarize(mean_germ = mean(prop_germ, na.rm = TRUE)) %>%
  ggplot(aes(x=timepoint, 
             y=mean_germ, 
             color=microbe_group))+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(
    title="Mean Germination Over Time for Each Microbe Group",
    x="Timepoint",
    y="Mean Proportion Germinated",
    color="Microbe Group")
```

#standard OD & all germination
- check how different OD changed germination
- suggest larger OD and longer time point have greater germination
```{r}
pollen_micriobe%>%
  ggplot(aes(x=factor(standardized_od_plate), 
             y=allgerm, 
             fill=microbe_group))+
  geom_boxplot()+
  theme_minimal()+
  facet_wrap(~timepoint)+
  labs(
    title="Effect of OD on Germination",
    x="Initial OD",
    y="All Germination")
```

