---
title: "Pollen Germination and Microbe Treatment"
author: "Chueh-Ju Kuo and Chelsea Ha"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library("ggthemes")
library("RColorBrewer")
library("paletteer")
library(ggplot2)
``` 

##Contents of the Project: 

##Todo List

1. Clean Data
    - make nice names with underscores 
    - change the ns (not significant) to NA so that we can use naniar to visualize missing data and also so that it         doesn't mess with our graphs.
    
2. Introduction - what is our research topic? List ideas below:
    - pollen germination and bursting across different microbial treatments and how this affects the microbial density
    
3. Methods - how will we analyze
    - use tidyverse to clean data and make graphs, use shiny to make an interactive app to explore the data
    
4. Visuals - graphs, what are our key variables?
    - microbe treatment, time point, proportion germinated, proportion burst, microbial density

5. Conclusion: 
    - summarize what we found and how it relates to the research question. 
    - future directions: what are some next steps we can take to further explore this topic?

6. Organize Repository (going off the example given, but we can refine)
    - datasets folder
    - background info folder - need to check if we have any
    - shiny app folder
    - exploring data
    - graphs
    
Data we used:



    
    
    
