---
title: "Pollen Germination and Microbe Treatment"
author: "Chueh-Ju Kuo and Chelsea Ha"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/users/hache/OneDrive/Documents/Desktop/BIS15L_W2026_Group1/Data")
```

## Load the libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library("ggthemes")
library("RColorBrewer")
library("paletteer")
library(ggplot2)
```

##Contents of the Project:

##Todo List

1.  Clean Data
    -   make nice names with underscores
    -   change the ns (not significant) to NA so that we can use naniar to visualize missing data and also so that it doesn't mess with our graphs.
2.  Introduction - what is our research topic? List ideas below:
    -   pollen germination and bursting across different microbial treatments and how this affects the microbial density
3.  Methods - how will we analyze
    -   use tidyverse to clean data and make graphs, use shiny to make an interactive app to explore the data
4.  Visuals - graphs, what are our key variables?
    -   microbe treatment, time point, proportion germinated, proportion burst, microbial density
5.  Conclusion:
    -   summarize what we found and how it relates to the research question.
    -   future directions: what are some next steps we can take to further explore this topic?
6.  Organize Repository (going off the example given, but we can refine)
    -   datasets folder
    -   background info folder - need to check if we have any
    -   shiny app folder
    -   exploring data
    -   graphs


##Data Exploration for each dataset:

##Data on pollen germination and bursting with different microbial treatments

```{r echo=TRUE}
pollen_microbe<-read_csv("pollen_microbe_data.csv")%>%
  clean_names()
```

Change the names of microbetreatment to microbe_treatment

```{r}
pollen_microbe_new <- pollen_microbe %>%
   rename(microbe_treatment = microbetreatment)
```

Change NA values to 0 in columns with NA

```{r}
pollen_microbe_new <- pollen_microbe_new %>%
  mutate(across(where(is.numeric), ~replace_na(., 0)))
```

```{r}
pollen_microbe_new %>% 
  distinct(microbe_group, microbe_treatment)
```

Arrange the data from the most germinated to the least germinated for each time point and microbe treatment

```{r}
pollen_microbe_new %>%
  select(timepoint, microbe_group, microbe_treatment, allgerm) %>%
  arrange(desc(allgerm))
```

```{r echo=FALSE}
pollen_microbe_new %>% 
  group_by(timepoint, microbe_group) %>%
  ggplot(aes(x = timepoint, y = allgerm, fill = microbe_group))+
  geom_col(position = "dodge")+
  theme_bw()
```
The acinetobacter has the largest amount of germination at each timepoint, followed by the control, then the yeast. The control pectin degrader was used to rule out the idea that simply breaking open the pollen (by degrading pectin in the inner layer of the pollen grain) would result in germination; we can see that the control pectin degrader has very low germination, so we can rule out that idea. The acinetobacter treatment has the highest germination, which suggests that the acinetobacter is inducing germination.

"Evidence presented here—including minimal effects of Pectobacterium, a known pectinase producer, and a lack of detected pectinase activity by Acinetobacter—suggest that pectinases alone are not responsible for the germination or bursting phenotype."


We can see if the type of microbial group has any correlation with the number of germinated pollen (pollen tubes are fully grown), the number of grains that burst (pollen tubes are not fully grown and the pollen grain is broken), and the number of grains that burst at the tip.
```{r echo=TRUE}
pollen_microbe_new %>%
  group_by(microbe_group, timepoint) %>%
  ggplot(aes(x = microbe_group, y = number_germ, fill = microbe_group))+
  geom_col(position = "dodge")+
  theme_bw()+
  labs(x = "Microbe Group", y = "# Germinated", fill = "Microbe Group", title="# Germinated Pollen Grains with Different Microbial Groups Over Time")+
  facet_wrap(~timepoint)+
  theme(axis.text.x = element_text(angle = 90))
```
```{r}
pollen_microbe_new %>%
  group_by(microbe_group, timepoint) %>%
  ggplot(aes(x = microbe_group, y = number_burst, fill = microbe_group))+
  geom_col(position = "dodge")+
  theme_bw()+
  labs(x = "Microbe Group", y = "# Burst", fill = "Microbe Group", title="# Burst Pollen Grains with Different Microbial Groups Over Time")+
  facet_wrap(~timepoint)+
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
pollen_microbe_new %>%
  group_by(microbe_group, timepoint) %>%
  ggplot(aes(x = microbe_group, y = number_tip_burst, fill = microbe_group))+
  geom_col(position = "dodge")+
  theme_bw()+
  labs(x = "Microbe Group", y = "# Tip Burst", fill = "Microbe Group", title="# Tip Burst Pollen Grains with Different Microbial Groups Over Time")+
  facet_wrap(~timepoint)+
  theme(axis.text.x = element_text(angle = 90))
```
From the three graphs, the acinetobacter increases total pollen bursting and tip bursting over time as well as in the burst before germination (no visible pollen tube). Using the number_burst variable, meaning this is before the pollen tube is fully grown, we can see that the acinetobacter has the highest number of burst pollen grains at each time point. This may mean that acinetobacter will induce burst-before-germination events, resulting in nutrients being released from the pollen grain before the pollen tube is fully grown.



##Data on proportion of grains germinated with different dilutions of A. pollinisSCC477 over time.

```{r}
dose_dependence <-read_csv("dose_dependence_data.csv")%>%
  clean_names() %>% 
  rename(microbe_treatment = microbetreatment)
```

```{r}
glimpse(dose_dependence)
```

In a separate assay, we used a dilution series of A.pollinis SCC477 to determine whether concentration of it had an effect of pollen germination and bursting over 24 h. Make a table
```{r}
dose_dependence %>%
  group_by(microbe_treatment, timepoint) %>%
  summarise(mean_germinated = mean(allgerm),
            mean_burst = mean(allburst)) %>% 
  arrange(desc(mean_germinated))
```

Make a density plot showing the germination overtime for 1:20 dilution of A. pollinis SCC477
```{r}
dose_dependence %>%
  filter(microbe_treatment == "(1:20)") %>%
  ggplot(aes(x = minutes/60, y = allgerm))+
  geom_col(position = "dodge")+
  theme_bw()+
  labs(x = "Hours", y = "# Germinated", title="# Germinated Pollen Grains with 1:20 Dilution of A. pollinis SCC477 Over Time")
```


##Data on microbial hemocytometer counts of A. pollinis, when provided different pollen treatments.

```{r}
a_pollinis <- read_csv("Apollinis_benefit_data.csv") %>%
  clean_names()
```

##Data on microbial hemocytometer counts of M. reukaufii, when provided different pollen treatments.

```{r}
yeast_benefit <- read_csv("yeast_benefit_data.csv") %>%
  clean_names()
```

##Data on germination and bursting after pollen (germinable or ungerminable) exposure to SCC477 (or no inoculation).

```{r}
germination_burst <- read_csv("germ_AN.csv") %>%
  clean_names() %>% 
  rename(pollen_treatment = pollentreatment)
```

##Data on protein in solution after pollen (germinable or ungerminable or no pollen) exposure to SCC477 (or no inoculation).

```{r}
protein_density <- read_csv("Lowry_data1.csv") %>%
  clean_names() %>% 
  rename(pollen_treatment = pollentreatment,
         microbe_treatment = microbetreatment)
```




##Questions: 

## 1) What is the average number of Pollen germinated with microbial species compared to other microbial groups? (make sure to label the 15 mintues, 45 minutes, and 90 minutes on the x-axis)

```{r}
pollen_microbe_new %>%
  group_by(microbe_group, timepoint) %>%
  summarise(mean_germinated = mean(allgerm)) %>% 
  ggplot(aes(x = timepoint/15, y = mean_germinated, fill = microbe_group))+
  geom_col(position = "dodge")+
  coord_flip()+
  labs(x = "Time Point", y = "Mean # Germinated", fill = "Microbe Group", title="Mean # Germinated Pollen Grains with Different Microbial Groups Over Time")
```
```{r}
glimpse(germination_burst)
```
make 2 tables for the GP and UP for each treatment
```{r}
germination_burst %>%
  filter(pollen_treatment == "GP") %>%
  group_by(treatment) %>%
  summarise(mean_germinated = mean(allgerm),
            mean_burst = mean(allburst)) %>% 
  arrange(desc(mean_germinated))
```
```{r}
germination_burst %>%
  filter(pollen_treatment == "UP") %>%
  group_by(treatment) %>%
  summarise(mean_germinated = mean(allgerm),
            mean_burst = mean(allburst)) %>% 
  arrange(desc(mean_germinated))
```


## 2) How does the percentage of pollen germinated change over time with different microbial species? (Relationship between pollen germination and time)
```{r}

```

## 3) How does the microbial density change over time with different pollen conditions? (Relationship between bacterial density and pollen germination)

```{r}

```


## 4) Microbial growth with different pollen conditions (A.pollinis vs. Yeast (known not to induce germination process))
```{r}
a_pollinis_wide <- a_pollinis %>%
  group_by(date) %>% 
  pivot_wider(
    names_from = tube,
    values_from = final_count_24h
  )
a_pollinis_wide
```

```{r}
a_pollinis_new <- a_pollinis %>%
  mutate(delta = final_count_24h - initial_count) %>% 
  arrange(desc(delta))
```

```{r}
a_pollinis_new %>%
  group_by(date, tube) %>%
  summarise(mean_delta = mean(delta, na.rm = TRUE)) %>%
  ggplot(aes(x = tube, y = mean_delta, fill = tube)) +
  geom_col() +   # geom_col is just geom_bar(stat = "identity")
  theme_bw()

```





## 5) Does pollen germination increase microbial growth?
```{r}
protein_density%>%
  group_by(pollen_treatment, microbe_treatment) %>%
  summarise(mean_protein = mean(protein)) %>%
  ggplot(aes(x = pollen_treatment, y = mean_protein, fill = microbe_treatment))+
  geom_col(position = "dodge")+
  theme_bw()+
  labs(x = "Pollen Treatment", y = "Mean Protein Concentration", fill = "Microbe Treatment", title="Mean Protein Concentration with Different Pollen and Microbe Treatments")
```






##Shiny App
```{r}
library(shiny)

ui <- fluidPage(
  
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```








