---
title: "Pollen Germination and Microbe Treatment"
author: "Chueh-Ju Kuo and Chelsea Ha"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: yes
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/users/hache/OneDrive/Documents/Desktop/BIS15L_W2026_Group1/Data")
```

## Load the libraries 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library("ggthemes")
library("RColorBrewer")
library("paletteer")
library(ggplot2)
``` 

##Contents of the Project: 


##Todo List

1. Clean Data
    - make nice names with underscores 
    - change the ns (not significant) to NA so that we can use naniar to visualize missing data and also so that it         doesn't mess with our graphs.
    
2. Introduction - what is our research topic? List ideas below:
    - pollen germination and bursting across different microbial treatments and how this affects the microbial density
    
3. Methods - how will we analyze
    - use tidyverse to clean data and make graphs, use shiny to make an interactive app to explore the data
    
4. Visuals - graphs, what are our key variables?
    - microbe treatment, time point, proportion germinated, proportion burst, microbial density

5. Conclusion: 
    - summarize what we found and how it relates to the research question. 
    - future directions: what are some next steps we can take to further explore this topic?

6. Organize Repository (going off the example given, but we can refine)
    - datasets folder
    - background info folder - need to check if we have any
    - shiny app folder
    - exploring data
    - graphs
    

##Insert all the Data:
```{r}
getwd()
```
```{r}
pollen_microbe<-read_csv("pollen_microbe_data.csv")%>%
  clean_names()
```
```{r}
dose_dependence<-read_csv("dose_dependence_data.csv")%>%
  clean_names()
```
```{r}
a_pollinis <- read_csv("Apollinis_benefit_data.csv") %>%
  clean_names()
```

```{r}
yeast_benefit <- read_csv("yeast_benefit_data.csv") %>%
  clean_names()
```
```{r}
germination_burst <- read_csv("germ_AN.csv") %>%
  clean_names()
```
```{r}
protein_density <- read_csv("Lowry_data1.csv") %>%
  clean_names()
```




## What is the percentage of Pollen germinated with microbial species compared to a control variable (yeast)?
```{r}

```


## How does the percentage of pollen germinated change over time with different microbial species? (Relationship between pollen germination and time)
```{r}

```

## How does the microbial density change over time with different pollen conditions? (Relationship between bacterial density and pollen germination)
```{r}

```

## Microbial growth with different pollen conditions (A.pollinis vs. Yeast (known not to induce germination process))
```{r}

```

## Does pollen germination increase microbial growth?   
```{r}
names(protein_density)
```
Change the names of pollentreatment to pollen_treatment and microbetreatment to microbe_treatment
```{r}
protein_density %>%
  rename(pollen_treatment = pollentreatment,
         microbe_treatment = microbetreatment)
```

